Test parsing of Proff.no table

### 2022-07-14 Test parsing of Proff table.Rmd
New and improved function. Hurrah.

## Function with libraries
```{r}
read_proff5 <- function(file, start, end){
  require(rvest)
  require(tidyverse)
  require(janitor)
  read_html(file, encoding = "UTF-8") |> 
  html_node("#inner-frame") |> 
  html_table() |> 
  select( # removing empty column
    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
  mutate( # removing whitespace
    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
  ) |> 
  filter( # removing duplicate rows
    !grepl("Lukk", REGNSKAPSPERIODE)
    ) |> 
  pivot_longer( # tidying data
    all_of(end):all_of(start), names_to = "year") |> # using all_of()
  mutate( # changing to real NAs and turning years into numbers
    value = na_if(value, "-"),
    year = as.integer(year)) |> 
  mutate( # adding currency column
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |> 
  fill( # filling the currency column
    valutakode, .direction = "updown"
    ) |> 
  filter( # removing dates and redundant currency
    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
      REGNSKAPSPERIODE != "Valutakode") |> 
  mutate( # removing whitespace in numbers
    value = str_replace_all(value, "\\s", "")
  ) |> 
  mutate( # turning into numbers
    value = as.numeric(value)
  ) |> 
  mutate( # removing years as values
    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
  ) |> 
    distinct() |> # removing duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
    clean_names() |> # tidying names
    arrange(desc(year)) # arranging by year
}

read_proff5("Links/Riksmaal.html", "2011", "2021")

@@

>>>>

## Function with libraries

- `file`: name and location of html file
- `start`: first available year as a string ("")
- `end`: last available year as a string ("")
```{r}
read_proff5 <- function(file, start, end){
  require(rvest)
  require(tidyverse)
  require(janitor)
  read_html(file, encoding = "UTF-8") |> 
  html_node("#inner-frame") |> 
  html_table() |> 
  select( # removing empty column
    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
  mutate( # removing whitespace
    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
  ) |> 
  filter( # removing duplicate rows
    !grepl("Lukk", REGNSKAPSPERIODE)
    ) |> 
  pivot_longer( # tidying data
    all_of(end):all_of(start), names_to = "year") |> # using all_of()
  mutate( # changing to real NAs and turning years into numbers
    value = na_if(value, "-"),
    year = as.integer(year)) |> 
  mutate( # adding currency column
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |> 
  fill( # filling the currency column
    valutakode, .direction = "updown"
    ) |> 
  filter( # removing dates and redundant currency
    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
      REGNSKAPSPERIODE != "Valutakode") |> 
  mutate( # removing whitespace in numbers
    value = str_replace_all(value, "\\s", "")
  ) |> 
  mutate( # turning into numbers
    value = as.numeric(value)
  ) |> 
  mutate( # removing years as values
    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
  ) |> 
    distinct() |> # removing duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
    clean_names() |> # tidying names
    arrange(desc(year)) # arranging by year
}

riksmaal <- read_proff5("Links/Riksmaal.html", "2011", "2021")
names(riksmaal)
summary(riksmaal)

rodekors <- read_proff5("Links/Kors.html", "2001", "2020")

rodekors
```

## New function - ffs
```{r}
read_proff6 <- function(file, start, end){
  require(rvest)
  require(tidyverse)
  require(janitor)
  read_html(file, encoding = "UTF-8") |> 
  html_node("#inner-frame") |> 
  html_table() |> 
  select( # removing empty column
    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
  mutate( # removing whitespace
    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
  ) |> 
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
  filter( # removing duplicate rows
    !grepl("Lukk", REGNSKAPSPERIODE)
    ) |> 
  pivot_longer( # tidying data
    all_of(end):all_of(start), names_to = "year") |> # using all_of()
  mutate( # changing to real NAs and turning years into numbers
    value = na_if(value, "-"),
    year = as.integer(year)) |> 
  mutate( # adding currency column
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |> 
  fill( # filling the currency column
    valutakode, .direction = "updown"
    ) |> 
  filter( # removing dates and redundant currency
    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
      REGNSKAPSPERIODE != "Valutakode") |> 
  mutate( # removing whitespace in numbers
    value = str_replace_all(value, "\\s", "")
  ) |> 
  mutate( # turning into numbers
    value = as.numeric(value)
  ) |> 
  mutate( # removing years as values
    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
  ) |> 
  mutate( # values no longer in 1000
      value = value * 1000
    ) |> 
  distinct() |> # removing duplicate rows
  pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
  clean_names() |> # tidying names
  select(-lederlonn_i_hele_1000,
           -resultatregnskap_i_hele_1000,
           -balanseregnskap_i_hele_1000) |> # removing headings
  rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |> 
  arrange(desc(year)) # arranging by year
}

read_proff6("Links/Kors.html", "2001", "2020")
read_proff6("Links/opera.html", "1998", "2021")
read_proff6("Links/Riksmaal.html", "2011", "2021")

@@

## Visualisations - 5th function
```{r}
riksmaal |> 
  ggplot(aes(year, arsresultat)) +
  geom_path()

riksmaal |> 
  ggplot(aes(year, sum_egenkapital_og_gjeld)) +
  geom_col()
